I don't know if projection is categorical. \\ Did \citealt{mandelkern-etal2020} discover that it is?

### 2023-11-29 writing/naturalness-paper/paper.tex
updated naturalness paper and analysis scripts

\author[$\circ$]{Judith Tonhauser}
\author[$\bullet$]{Judith Degen}
\affil[$\circ$]{University of Stuttgart}
\affil[$\bullet$]{Stanford University}

@@

\maketitle

@@

Presuppositions are taken to typically project out of entailment-canceling environments like the scope of negation (e.g., \citealt{ccmg90}). While projection is often characterized as categorical (that is, content either typically projects or not), there is mounting empirical evidence that projection is gradient, with content being more or less projective (e.g., \citealt{karttunen71b,xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}). \citealt{mandelkern-etal2020} critically evaluated the inference rating measures on which this empirical evidence is based and claimed that a different measure, namely naturalness ratings in explicit ignorance contexts, provides support for categorical projection and categorically distinguishes presuppositions from nonpresuppositions. This paper presents the results of an experiment designed to investigate \citepos{mandelkern-etal2020} claim for factive predicates (presumed presupposition triggers) and nonfactive ones (nontriggers). The results do not support \citepos{mandelkern-etal2020} claim and rather align with \citepos{degen-tonhauser-language} result that there is no evidence for a categorical distinction between factive and nonfactive predicates.

@@

@@

@@

Many theoretical works assume that projection is a categorical property of utterance content, that is, content either typically projects (in which case it is a presupposition) or it does not (in which case it is not a presupposition). There is, however, mounting empirical evidence that projection is not a categorical property of  content but rather a gradient one, with content being more or less projective. The measure used to establish this evidence are inference rating measures: Such measures involve presenting participants with an utterance (like (\ref{first}a)) and contents (like that Julian dances salsa), and asking for ratings that are taken to reveal the existence or strength of the inference to the content. \citealt{karttunen71b}, for instance, suggested that the CCs of {\em regret} and {\em discover}, which he took to be presuppositions, exhibit projection variability in examples like (\ref{kart}). Specifically, he suggested that a speaker who utters (\ref{kart}a) believes that the addressee has not told the truth, whereas a speaker who utters (\ref{kart}b) ``is not sure about the truth of the complement'' (p.63).

\begin{exe}
\ex\label{kart}
\begin{xlist}
\ex Did you regret that you had not told the truth?
\ex Did you discover that you had not told the truth? \hfill (\citealt[63]{karttunen71b})
\end{xlist}
\end{exe}
More recently, experimental investigations have also observed projection variability among presuppositions (e.g., \citealt*{xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}). \citealt{tbd-variability}, for instance, observed in their Exps.~1 that the CC of {\em know} is more projective than that of {\em discover}, and the CC of {\em discover} is more projective than that of {\em reveal}. These results led \citealt[498]{tbd-variability} to propose that ``projectivity [\ldots is] a gradient property of content, rather than a binary, categorical one''. Further support for this proposal comes from  investigations of both factive and nonfactive predicates in \citealt{degen-tonhauser-language}. In a series of experiments and analyses of existing datasets, this work not only found projection variability between the CCs of factive predicates, but also no empirical support for a categorical distinction in the projection of the CCs of factive and nonfactive predicates. This is because the CCs of both factive and nonfactive predicates were projective to varying degrees compared to nonprojective main clause content. A sample illustration of these results is given in Fig.~\ref{fig:dt1a} from their Exp.~1a, which shows the mean projection ratings of 5 factive predicates (in \color{orange}orange\color{black}) and 15 nonfactive predicates (in \color{green}green\color{black}).

@@

\citealt[\S6.2]{mandelkern-etal2020} suggested that naturalness ratings of utterances in explicit ignorance contexts are better suited to distinguish presuppositions from nonpresuppositional projection inferences.\footnote{\citealt{mandelkern-etal2020} refer to the measure as an acceptability rating measure, but since participants were asked to rate the naturalness of utterances, we refer to it as a naturalness rating measure.} On this measure, participants are presented with utterances in which a speaker states that they are ignorant about a content, as in (\ref{mtrig}a), and then follows up with an utterance in which the content occurs embedded in an entailment-canceling environment, such as the antecedent of a conditional in (\ref{mtrig}c). \citealt{mandelkern-etal2020} assumed that if the content is a presupposition (such as the pre-state content of {\em stop} in (\ref{mtrig}c)), the utterance is rated as unnatural in the explicit ignorance context because presuppositions are contents that the speaker believes to be true. That is, in such constellations,  participants ``will not have an alternative to interpreting the presupposition at the utterance level, and in turn seeing the utterance as infelicitous and the speaker as incoherent'' (p.497).  If, on the other hand, the content is not a presupposition (such as the pre-state content of {\em now frown on} in (\ref{mtrig}c)), the utterance is assumed to be natural: Even if participants might, in a different context, draw a projection inference ``for any of a variety of pragmatic reasons short of entailment or presupposition, [\ldots] they will tend to relinquish that inclination when there is pragmatic pressure to do so'' as in the explicit ignorance contexts (p.497). Both presuppositions and nonpresuppositions were expected to be judged as natural in the support contexts, which entail the relevant content, illustrated in (\ref{mtrig}b).

@@

In \citepos{mandelkern-etal2020} Exp.~3, which used the proposed measure, items like (\ref{mtrig}) were control items. Mandelkern and colleagues found that the naturalness ratings for nontriggers were relatively high in both contexts (with means just below 5, where 1 meant ``completely unnatural'' and 7 meant ``completely natural'' on their 7-point Likert scale). The presupposition triggers, on the other hand, had a lower mean naturalness rating in the explicit ignorance context (mean around 3) than in the support context (mean at about 5.3). Mandelkern and colleagues took this to ``establish[] the effectiveness of the methodology'' (p.492). They proposed (p.497):

@@

However, the empirical evidence in \citealt{mandelkern-etal2020} in support of the claim projection is categorical is quite weak. This is primarily because the goal of their work was not to investigate this claim, but rather whether presupposition projection in conjunctions is symmetric. It is also weak, however, because the data provided in Fig.~8 in their Appendix 3 do not appear to support a categorical distinction between the four presupposition triggers ({\em be aware, be happy, stop, continue}) and the four nontriggers ({\em now frown on, enjoy, hope, be sure}) that were investigated. First, considering just the mean naturalness ratings in the explicit ignorance context, it appears that the ratings for the presuppositions and nonpresuppositions overlap: The ratings for the presuppositions ranged from 1 (for the pre-state of {\em continue}) to about 3.5 (for the CC of {\em be aware}), and the ratings for the nonpresuppositions range from just below 3 (for the CC of {\em hope}) to 4.5 (for the CC of {\em be aware}). It is therefore not clear that naturalness ratings in explicit ignorance contexts provide evidence for a categorical distinction between the four presuppositions and the four nonpresuppositions investigated in \citealt{mandelkern-etal2020}. Second, there is also no consistent effect of context. Whereas the difference between the mean naturalness ratings in the explicit ignorance and support contexts was quite large for the (presuppositional) pre-state of {\em continue}, the difference was much less pronounced for the (presuppositional) CC of {\em be aware}, where it was roughly comparable to that of the (nonpresuppositional) pre-state of {\em frown on}. Thus, it is not clear that comparisons of ratings in support and explicit ignorance contexts provides empirical support for a categorical projection either.

@@

Another issue with \citepos{mandelkern-etal2020} claim is that there is disagreement in the literature about the naturalness of presuppositions in explicit ignorance contexts. Whereas Mandelkern and colleagues assumed that presuppositions are unnatural in explicit ignorance contexts, \citealt{simons01} and \citealt{abusch10} assumed that only `hard' (nondefeasible) presuppositions are, but that `soft' (defeasible) ones are natural. For instance,  the examples with {\em stop}, {\em discover}, and {\em win} in (\ref{eic2}) were assumed to be natural in explicit ignorance contexts, and the examples with {\em again}, the {\em it-}cleft and {\em too} in (\ref{eic3}) unnatural.\footnote{\citealt{abusch10} marked (\ref{eic3}b) and (\ref{eic3}c) with `??' instead of `$\#$'.}

@@

\ex \infelic I have no idea whether anyone read that letter. But if it is John
who read it, let's ask him to be discreet about the content. \hfill (\citealt[40]{abusch10})

\ex \infelic I have no idea whether John read that proposal. But if Bill read it too, let's ask them to confer and simply give us a yes-no response. \hfill (\citealt[40]{abusch10})
\end{xlist}
\end{exe}

This paper presents the results of an experiment designed to investigate \citepos{mandelkern-etal2020} claim that the measure they proposed provides evidence that projection is categorical and distinguishes presuppositions from nonpresuppositions.\footnote{Link to experiment, materials, and analysis scripts: \url{xxx} {\bf DO THIS AFTER FINALIZING SCRIPTS}}  To allow for comparison with the results of \citealt{degen-tonhauser-language}, the CCs of the same 20 (non)factive clause-embedding predicates (see Fig.~\ref{fig:dt1a}) were investigated. The experiment also included control stimuli with {\em stop, continue, again, too, also}, and an {\em it-}cleft to allow for comparison with the results of \citealt{mandelkern-etal2020} and the assumptions made in \citealt{simons01} and \citealt{abusch10}.

We slightly changed the design of \citepos{mandelkern-etal2020} Exp.~3 to allow for comparison with the result of \citealt{degen-tonhauser-openmind} that the projection of content is sensitive to the content's prior probability. Specifically, what \citealt{degen-tonhauser-openmind} found was that the higher the prior probability of a content, the stronger the projection inference. For instance, a speaker's commitment to the CC of (\ref{prior}c), that Julian dances salsa, was stronger in the context in (\ref{prior}b), where the CC has a higher prior probability, than in the context in (\ref{prior}a), where the CC has a lower prior probability.

@@

To investigate \citepos{mandelkern-etal2020} claim that their naturalness rating measure provides support for a categorical distinction between presuppositions and nonpresuppositions, participants read two-sentence discourses consisting of a declarative (which provided the context) and an interrogative (which realized a (non)factive predicate), and rated the naturalness of the interrogative in the context of the declarative.

@@

We recruited 425 participants on Prolific. Due to a server error, the data of only 398 participants was recorded (ages: 19-73, mean: 40.8; 187 women, 201 men, 8 non-categorical, 2 preferred to not disclose). The recruited participants were required to live in the USA, to speak English as their first language, to have completed at least 100 tasks, and to have an approval rating of at least 99\%. The median time spent on the task was 6:24 minutes. Participants were paid \$1.78, corresponding to an hourly pay of \$16.6.

@@

Participants read two-sentence discourses consisting of a declarative followed by an interrogative, as shown in (\ref{sample}). In the target stimuli, the interrogatives combined the 20 (non)factive predicates of \citealt{degen-tonhauser-language} (see Fig.~\ref{fig:dt1a}) with the 20 complement clauses of \citealt{degen-tonhauser-language}, for a total of 400 interrogatives. The preceding declaratives implemented a three-level context condition: In the `explicit ignorance' context (\ref{sample}a), the declarative sentence conveyed the speaker's ignorance about the CC. In the `lower prior probability' context (\ref{sample}b), the CC had a comparatively lower prior probability, whereas it had a comparatively higher prior probability in the `higher prior probability' context (\ref{sample}c). The two contexts for each of the 20 complement clauses that were combined with the 20 predicates were normed in \citealt{degen-tonhauser-openmind}. See Supplement \ref{a:clauses} for the full set of 20 complement clauses and the two contexts for each complement clause.

@@

In addition to the 1,200 target stimuli (400 interrogatives $\times$ 3 contexts), the materials also included the six control stimuli shown in (\ref{filler}). The interrogatives of the control stimuli featured six expressions typically analyzed as presupposition triggers, namely {\em stop, continue, again, too, also}, and an {\em it-}cleft. The preceding declaratives conveyed the speaker's explicit ignorance about the presuppositions associated with these expressions.\footnote{The contents investigated for {\em too} and {\em also} in (\ref{filler}d) and (\ref{filler}e) are that Ann plays an instrument other than the flute and that Svenja plays a sport other than soccer. This interpretation arises if {\em Ann} and {\em Svenja} associate with {\em too} and {\em also}, respectively. While prosody was not controlled for, this focus association is made plausible by the preceding explicit ignorance statements, which evoke the question of whether Ann plays an instrument and whether Svenja plays any sport, respectively.} Two of these expressions, namely {\em stop} and {\em continue}, were target expressions in \citepos{mandelkern-etal2020} Exp.~3, where they were assumed to be unnatural in explicit ignorance contexts. This is in contrast to \citealt{simons01}, who took {\em stop} to be a soft trigger that is natural in explicit ignorance contexts. Three expressions were assumed to be unnatural in explicit ignorance contexts in \citealt{simons01} ({\em again}) and \citealt{abusch10} ({\em too}, {\em it-}cleft). The control stimuli were included for comparison to the target stimuli, to the claims of \citealt{simons01} and \citealt{abusch10}, and to the results of \citealt{mandelkern-etal2020}.

@@

A random set of 30 stimuli was created for each participant. Each set contained 20 target stimuli in which each of the 20 complement clauses was paired with a unique clause-embedding predicate. Twelve of the target stimuli were presented in the explicit ignorance context, and the other eight in a low or a higher prior probability context (four each). Each participants' set of 30 stimuli contained the same six control stimuli and the same four filler stimuli. Each of the 30 stimuli were presented as utterances by a unique named speaker. Trial order was randomized for each participant.

@@

Participants were instructed to rate ``how natural the question sounds in the context of the statement''. As shown in Figure \ref{f:trials}, they were asked give their rating on a slider from `totally unnatural' (coded as 0) to `totally natural' (coded as 1). The experiment began with four practice trials to familiarize participants with the task (see Supplement \ref{a:fillerPractice} for details). After rating the 30 trials, participants filled out a short optional demographic survey. To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.

@@

We excluded the data of five participants who did not self-identify as native speakers of American English as well as the data of 23 participants whose mean response to the filler stimuli (expected to be natural) was more than 2 sd below the group mean.\footnote{Contrary to what was preregistered, one of the four filler stimuli was not used to exclude participants because it received a mean rating of only .5. See Supplement \ref{a:fillerPractice} for details.} The data of 370 participants entered into the analysis (ages: 19-80, mean: 40.7; 175 women, 185 male, 8 non-categorical, 2 preferred to not disclose). Each of the 20 predicates received at least 200 ratings in the explicit ignorance context (mean: 222 ratings), at least 59 ratings in the lower prior probability context (mean: 74), and at least 61 ratings in the higher prior probability context (mean: 74). The six control stimuli received 370 ratings each.

@@

We first address the question of whether naturalness ratings in explicit ignorance contexts provide empirical evidence for a categorical difference between presuppositions and nonpresuppositions (\S\ref{s:analysis1}). We then consider whether a comparison of naturalness ratings in the three contexts provides such evidence (\S\ref{s:analysis2}).

@@

These observations were confirmed by a posthoc pairwise comparison of the naturalness ratings of the contents associated with the six controls and 20 target expressions in the explicit ignorance context using the `emmeans' package (\citealt{emmeans}) in R (\citealt{r}). The input to the pairwise comparison was a Bayesian mixed-effects beta regression model with weakly informative priors that was fitted using the `brms' package (\citealt{buerkner2017}). The model predicted naturalness ratings from a fixed effect of expression (with treatment coding and `continue' as reference level) and included random by-participant and by-item intercepts (where an item is a complement clause). The output of the pairwise comparison are 95\% highest density intervals of estimated marginal mean differences between each of the 26 contents under investigation. We assume that two contents differ in their naturalness in the explicit ignorance context if the 95\% highest density interval does not include 0.

As shown in Table \ref{t:pairwise},\footnote{The full model output is available here: \url{LINK TO TABLE IN REPO} {\bf ADD THIS AFTER FINALIZING SCRIPTS}.} the pairwise comparison suggests differences between the naturalness ratings of the CCs of the five factive predicates: The ratings of {\em be annoyed} are lower than those of {\em know}, which in turn are lower than those of {\em see} and {\em discover}, which in turn are lower than those of {\em reveal}. There is also variation between the factive predicates and the presuppositional controls: The CCs of each of the five purportedly factive predicates are more natural in explicit ignorance contexts than the presuppositions contributed by {\em too} and {\em also}. Furthermore, the naturalness ratings for the CCs of the five factive predicates are indistinguishable from different sets of expressions: The naturalness ratings for {\em be annoyed} do not differ from those of {\em stop} (but are higher than those the prejacents of the controls {\em continue, too}, and {\em also}), the naturalness ratings for {\em know} do not differ from those of the {\em it-}cleft, the naturalness ratings for {\em see} and {\em discover} do not differ from those of the nonfactives {\em confess, inform}, and {\em acknowledge}, and those of {\em reveal} do not differ from those of the nonfactives {\em admit, hear}, and {\em prove}. These results suggest that naturalness ratings in explicit ignorance contexts do not provide empirical support for a categorical distinction between factive and nonfactive predicates, or between presuppositions and nonpresuppositions.

>>>>

\author[$\circ$]{Judith Tonhauser}
\author[$\bullet$]{Judith Degen}
\affil[$\circ$]{University of Stuttgart, Department of Linguistics, Stuttgart, Germany, judith.tonhauser@ling.uni-stuttgart.de (corresponding author)}
\affil[$\bullet$]{Stanford University, Department of Linguistics, Stanford, USA, jdegen@stanford.edu}

@@

\maketitle
\thispagestyle{empty}

@@

Presuppositions are taken to typically project out of entailment-canceling environments like the scope of negation (e.g., \citealt{ccmg90}). While projection is often characterized as categorical (that is, content either typically projects or not), there is mounting empirical evidence that projection is gradient, with content being more or less projective (e.g., \citealt{karttunen71b,xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}). \citealt{mandelkern-etal2020} critically evaluated the inference rating measures on which this empirical evidence is based and claimed that a different measure, namely naturalness ratings in explicit ignorance contexts, provides support for categorical projection and distinguishes presuppositions from nonpresuppositions. This paper presents the results of an experiment designed to investigate \citepos{mandelkern-etal2020} claim for factive predicates (presumed presupposition triggers) and nonfactive ones (nontriggers). The results do not support \citepos{mandelkern-etal2020} claim and rather align with \citepos{degen-tonhauser-language} result that there is no evidence for a categorical distinction between factive and nonfactive predicates.

@@

\noindent
{\bf Funding:} The experiment was funded by Judith Tonhauser's research funds.

\noindent
{\bf Competing interests:} The authors have no competing interests to declare.

\noindent
{\bf Ethics approval and informed consent:} The experiment was approved by the ethics review committee of the University of Stuttgart. Informed consent was obtained from the participants.

\noindent
{\bf Availability of materials, data, and code:} See footnote \ref{f:github}.

\noindent
{\bf Author contributions:} Both authors contributed to the conception and design of the experiment. Material preparation, data collection and analysis were performed by both authors. The first draft of the manuscript was written by Judith Tonhauser. Both authors commented on previous versions of the manuscript. Both authors read and approved the final manuscript.

@@

\clearpage
\pagenumbering{arabic}

\begin{center}
{\LARGE I don't know if projection is categorical. \\[.2cm] Did \citealt{mandelkern-etal2020} discover that it is?}	
\end{center}

Presuppositions are taken to typically project out of entailment-canceling environments like the scope of negation (e.g., \citealt{ccmg90}). While projection is often characterized as categorical (that is, content either typically projects or not), there is mounting empirical evidence that projection is gradient, with content being more or less projective (e.g., \citealt{karttunen71b,xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}). \citealt{mandelkern-etal2020} critically evaluated the inference rating measures on which this empirical evidence is based and claimed that a different measure, namely naturalness ratings in explicit ignorance contexts, provides support for categorical projection and distinguishes presuppositions from nonpresuppositions. This paper presents the results of an experiment designed to investigate \citepos{mandelkern-etal2020} claim for factive predicates (presumed presupposition triggers) and nonfactive ones (nontriggers). The results do not support \citepos{mandelkern-etal2020} claim and rather align with \citepos{degen-tonhauser-language} result that there is no evidence for a categorical distinction between factive and nonfactive predicates.

@@

Many theoretical works assume that projection is a categorical property of  content, that is, content either typically projects (in which case it is a presupposition) or it does not (in which case it is not a presupposition). There is, however, mounting empirical evidence that projection is not a categorical property of  content but rather a gradient one, with content being more or less projective. The measure used to establish this evidence are inference rating measures: Such measures involve presenting participants with an utterance (like (\ref{first}a)) and contents (like that Julian dances salsa), and asking for ratings that are taken to reveal the existence or strength of the inference to the content. \citealt{karttunen71b}, for instance, suggested that the CCs of {\em regret} and {\em discover}, which he took to be presuppositions, exhibit projection variability in examples like (\ref{kart}). Specifically, he suggested that a speaker who utters (\ref{kart}a) believes that the addressee has not told the truth, whereas a speaker who utters (\ref{kart}b) ``is not sure about the truth of the complement'' (p.63).

\begin{exe}
\ex\label{kart}
\begin{xlist}
\ex Did you regret that you had not told the truth?
\ex Did you discover that you had not told the truth? \hfill (\citealt[63]{karttunen71b})
\end{xlist}
\end{exe}
More recently, experimental investigations also observed projection variability among presuppositions (e.g., \citealt*{xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}). \citealt{tbd-variability}, for instance, observed in their Exps.~1 that the CC of {\em know} is more projective than that of {\em discover}, which is more projective than that of {\em reveal}. These results led \citealt[498]{tbd-variability} to propose that ``projectivity [\ldots is] a gradient property of content, rather than a binary, categorical one''. Further support for this proposal comes from  investigations of both factive and nonfactive predicates in \citealt{degen-tonhauser-language}. In a series of experiments and analyses of existing datasets, this work not only found projection variability between the CCs of factive predicates, but also no empirical support for a categorical distinction in the projection of the CCs of factive and nonfactive predicates. This is because the CCs of both factive and nonfactive predicates were projective to varying degrees compared to nonprojective main clause content. A sample illustration of these results is in Fig.~\ref{fig:dt1a} from their Exp.~1a, which shows the mean projection ratings of 5 factive predicates (in \color{orange}orange\color{black}) and 15 nonfactive predicates (in \color{green}green\color{black}).\footnote{The data from \citepos{degen-tonhauser-language} Exp.~1a were accessed here: \url{https://github.com/judith-tonhauser/projective-probability/tree/master/results/5-projectivity-no-fact}. The script that was used to create Fig.~\ref{fig:dt1a} can be found in the GitHub repository linked in footnote \ref{f:github}.}

@@

\citealt[\S6.2]{mandelkern-etal2020} suggested that naturalness ratings of utterances in explicit ignorance contexts are better suited to distinguish presuppositions from nonpresuppositional projection inferences.\footnote{\citealt{mandelkern-etal2020} refer to the measure as an acceptability rating measure, but since participants were asked to rate the naturalness of utterances, we refer to it as a naturalness rating measure.} On this measure, participants are presented with utterances in which a speaker states that they are ignorant about a content, as in (\ref{mtrig}a), and then follows up with an utterance in which the content occurs embedded in an entailment-canceling environment, such as the antecedent of a conditional in (\ref{mtrig}c). \citealt{mandelkern-etal2020} assumed that if the content is a presupposition (such as the pre-state content of {\em stop} in (\ref{mtrig}c)), the utterance is rated as unnatural in the explicit ignorance context because presuppositions are contents that the speaker believes to be true. That is, in such constellations,  participants ``will not have an alternative to interpreting the presupposition at the utterance level, and in turn seeing the utterance as infelicitous and the speaker as incoherent'' (p.497).  If, on the other hand, the content is not a presupposition (such as the pre-state content of {\em now frown on} in (\ref{mtrig}c)), the utterance is assumed to be natural: Even if participants might, in a different context, draw a projection inference ``for any of a variety of pragmatic reasons short of entailment or presupposition, [\ldots] they will tend to relinquish that inclination when there is pragmatic pressure to do so'', as in the explicit ignorance contexts (p.497). Both presupposition triggers and nontriggers were expected to be judged as natural in the support context, which entails the relevant content, illustrated in (\ref{mtrig}b).

@@

In \citepos{mandelkern-etal2020} Exp.~3, which used the proposed measure, items like (\ref{mtrig}) were control items. Mandelkern and colleagues found that the naturalness ratings for nontriggers were relatively high in both contexts (with means just below 5 on their 7-point Likert scale, where 1 meant ``completely unnatural'' and 7 meant ``completely natural''). The presupposition triggers, on the other hand, had a lower mean naturalness rating in the explicit ignorance context (mean around 3) than in the support context (mean at about 5.3). Mandelkern and colleagues took this to ``establish[] the effectiveness of the methodology'' (p.492). They proposed (p.497):

@@

However, the empirical evidence in \citealt{mandelkern-etal2020} in support of the claim projection is categorical is quite weak. This is primarily because the goal of their work was not to investigate this claim, but rather whether presupposition projection in conjunctions is symmetric. It is also weak, however, because only four presupposition triggers ({\em be aware, be happy, stop, continue}) and four nontriggers ({\em now frown on, enjoy, hope, be sure}) were investigated, and the results (provided in Fig.~8 in their Appendix 3) do not appear to support a categorical distinction between the four triggers and the four nontriggers.  First, considering just the mean naturalness ratings in the explicit ignorance context, it appears that the ratings for the presuppositions and nonpresuppositions overlap: The ratings for the presuppositions ranged from 1 (for the pre-state of {\em continue}) to about 3.5 (for the CC of {\em be aware}), and the ratings for the nonpresuppositions range from just below 3 (for the CC of {\em hope}) to 4.5 (for the CC of {\em be aware}). It is therefore not clear that naturalness ratings in explicit ignorance contexts provide evidence for a categorical distinction between the four presuppositions and the four nonpresuppositions investigated in \citealt{mandelkern-etal2020}. Second, there is also no consistent effect of context. Whereas the difference between the mean naturalness ratings in the explicit ignorance and support contexts was quite large for the (presuppositional) pre-state of {\em continue}, the difference was much less pronounced for the (presuppositional) CC of {\em be aware}, where it was roughly comparable to that of the (nonpresuppositional) pre-state of {\em frown on}. Thus, it is not clear that comparisons of ratings in support and explicit ignorance contexts provides empirical support for a categorical projection either.

@@

Another issue with \citepos{mandelkern-etal2020} claim is that there is disagreement in the literature about the naturalness of presuppositions in explicit ignorance contexts. Whereas Mandelkern and colleagues assumed that presuppositions are unnatural in explicit ignorance contexts, \citealt{simons01} and \citealt{abusch10} assumed that only `hard' (nondefeasible) presuppositions are, but that `soft' (defeasible) ones are natural. For instance,  the examples with {\em stop}, {\em discover}, and {\em win} in (\ref{eic2}) were assumed to be natural in explicit ignorance contexts, and the examples with {\em again}, the {\em it-}cleft and {\em too} in (\ref{eic3}) unnatural.%\footnote{\citealt{abusch10} marked (\ref{eic3}b) and (\ref{eic3}c) with `??' instead of `$\#$'.}

@@

\ex \verymarginal I have no idea whether anyone read that letter. But if it is John
who read it, let's ask him to be discreet about the content. \hfill (\citealt[40]{abusch10})

\ex \verymarginal I have no idea whether John read that proposal. But if Bill read it too, let's ask them to confer and simply give us a yes-no response. \hfill (\citealt[40]{abusch10})
\end{xlist}
\end{exe}

This paper presents the results of an experiment designed to investigate \citepos{mandelkern-etal2020} claim that the measure they proposed provides evidence that projection is categorical and distinguishes presuppositions from nonpresuppositions.\footnote{\label{f:github}The experiment was approved by the ethics review committee of the [name of university suppressed]. The experiment, materials, data, and analysis scripts can be accessed here:  \url{xxx} {\bf DO THIS AFTER FINALIZING SCRIPTS}}  To allow for comparison with the results of \citealt{degen-tonhauser-language}, the CCs of the same 20 (non)factive clause-embedding predicates (see Fig.~\ref{fig:dt1a}) were investigated. The experiment also included control stimuli with {\em stop, continue, again, too, also}, and an {\em it-}cleft to allow for comparison with the results of \citealt{mandelkern-etal2020} and the assumptions made in \citealt{simons01} and \citealt{abusch10}.

The design of \citepos{mandelkern-etal2020} Exp.~3 was slightly changed to allow for comparison with the result of \citealt{degen-tonhauser-openmind} that the projection of content is sensitive to the content's prior probability. Specifically, what \citealt{degen-tonhauser-openmind} found was that the higher the prior probability of a content, the stronger the projection inference. For instance, a speaker's commitment to the CC of (\ref{prior}c), that Julian dances salsa, was stronger in the context in (\ref{prior}b), where the CC has a higher prior probability, than in the context in (\ref{prior}a), where the CC has a lower prior probability.

@@

To investigate \citepos{mandelkern-etal2020} claim that their naturalness rating measure provides evidence for categorical projection, participants read two-sentence discourses consisting of a declarative (which provided the context) and an interrogative (which realized a factive or nonfactive predicate), and rated the naturalness of the interrogative in the context of the declarative.

@@

We recruited 425 participants on Prolific. Due to a programming error, the data of only 398 participants was recorded (ages: 19-73, mean: 40.8; 187 women, 201 men, 8 non-categorical, 2 preferred to not disclose). The recruited participants were required to live in the USA, to speak English as their first language, to have completed at least 100 tasks, and to have an approval rating of at least 99\%. The median time spent on the task was 6:24 minutes. Participants were paid \$1.78, corresponding to an hourly pay of \$16.6.

@@

Participants read two-sentence discourses consisting of a declarative followed by an interrogative, as shown in (\ref{sample}). In the target stimuli, the interrogatives combined the 20 (non)factive predicates of \citealt{degen-tonhauser-language} (see Fig.~\ref{fig:dt1a}) with the 20 complement clauses of \citealt{degen-tonhauser-language}, for a total of 400 interrogatives. The preceding declaratives implemented a three-level context condition: In the `explicit ignorance' context (\ref{sample}a), the declarative sentence conveyed the speaker's ignorance about the CC. In the `lower prior probability' context (\ref{sample}b), the CC had a comparatively lower prior probability, whereas it had a comparatively higher prior probability in the `higher prior probability' context (\ref{sample}c). The two contexts for each of the 20 complement clauses were normed in \citealt{degen-tonhauser-openmind}. See Supplement \ref{a:clauses} for the full set of 20 complement clauses and the two contexts for each complement clause.

@@

In addition to the 1,200 target stimuli (400 interrogatives $\times$ 3 contexts), the materials also included the six control stimuli in (\ref{filler}). The interrogatives of the control stimuli featured six expressions typically analyzed as presupposition triggers, namely {\em stop, continue, again, too, also}, and an {\em it-}cleft. The preceding declaratives conveyed the speaker's explicit ignorance about the presuppositions associated with these expressions.\footnote{The contents investigated for {\em too} and {\em also} in (\ref{filler}d) and (\ref{filler}e) are that Ann plays an instrument other than the flute and that Svenja plays a sport other than soccer. This interpretation arises if {\em too} and {\em also} associate with {\em the flute} and {\em soccer}, respectively. While prosody was not controlled for, this focus association is made plausible by the preceding explicit ignorance statements, which evoke the question of whether Ann plays any instrument and whether Svenja plays any sport, respectively.} Two of these expressions, namely {\em stop} and {\em continue}, were target expressions in \citepos{mandelkern-etal2020} Exp.~3, where they were assumed to be unnatural in explicit ignorance contexts. This is in contrast to \citealt{simons01}, who took {\em stop} to be a soft trigger that is natural in explicit ignorance contexts. Three expressions were assumed to be unnatural in explicit ignorance contexts in \citealt{simons01} ({\em again}) and \citealt{abusch10} ({\em too}, {\em it-}cleft). The control stimuli were included for comparison to the target stimuli, to the claims of \citealt{simons01} and \citealt{abusch10}, and to the results of \citealt{mandelkern-etal2020}.

@@

A random set of 30 stimuli was created for each participant. Each set contained 20 target stimuli in which each of the 20 complement clauses was paired with a unique clause-embedding predicate. Twelve of the target stimuli were presented in the explicit ignorance context, and the other eight in a low or a higher prior probability context (four each). Each participants' set also contained the same six control stimuli and the same four filler stimuli. Each of the 30 stimuli were presented as utterances by a unique named speaker. Trial order was randomized for each participant.

@@

Participants were instructed to rate how natural the question sounds in the context of the statement. As shown in Figure \ref{f:trials}, they were asked give their rating on a slider from `totally unnatural' (coded as 0) to `totally natural' (coded as 1). The experiment began with four practice trials to familiarize participants with the task (see Supplement \ref{a:fillerPractice} for details). After rating the 30 trials, participants filled out a short optional demographic survey. To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.

@@

We excluded the data of 5 participants who did not self-identify as native speakers of American English and that of 23 participants whose mean response to the fillers (expected to be natural) was more than 2 sd below the group mean.\footnote{Contrary to what was planned, one of the four filler stimuli was not used to exclude participants' data because it received a mean rating of only .5. See Supplement \ref{a:fillerPractice} for details.} The data of 370 participants entered into the analysis (ages: 19-80, mean: 40.7; 175 women, 185 men, 8 nonbinary, 2 did not disclose). The 20 predicates each received at least 200 ratings in the explicit ignorance context (mean: 222 ratings), at least 59 ratings in the lower prior probability context (mean: 74), and at least 61 ratings in the higher prior probability context (mean: 74). The six controls received 370 ratings each.

@@

We first address the question of whether naturalness ratings in explicit ignorance contexts provide empirical evidence for categorical projection (\S\ref{s:analysis1}). We then consider whether a comparison of naturalness ratings in the three contexts provides such evidence (\S\ref{s:analysis2}).

@@

These observations were confirmed by a posthoc pairwise comparison of the naturalness ratings of the contents associated with the six controls and 20 target expressions in the explicit ignorance context using the `emmeans' package (\citealt{emmeans}) in R (\citealt{r}). The input to the pairwise comparison was a Bayesian mixed-effects beta regression model with weakly informative priors that was fitted using the `brms' package (\citealt{buerkner2017}). The model predicted naturalness ratings from a fixed effect of expression (with treatment coding and `continue' as reference level) and included random by-participant and by-item intercepts (where an item is a complement clause). The output of the pairwise comparison are 95\% highest density intervals of estimated marginal mean differences between each of the 26 contents under investigation. We assume that two contents differ in their naturalness in the explicit ignorance context if this 95\% highest density interval does not include 0.
